---
title: "Pre-processing of Dataset"
author: "Noel C. Sieras"
date: "2022-12-16"
output: pdf_document
---
## Preliminaries
This will prevent some errors in loading some of the chunks and laoding of the dataset. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'D:/FilesWorkOn&Saved/PhDStat/@MSUIIT/SY20222023/01FirstSemester/STT225_StatisticalComputing1/04FinalProject/02Dataset')
```
\
## Loading of packages
This package `tidyverse` will help the loading of the packeges needed for the pre=processing of data. The package `bestNormalize` will be used for normalizing the given dataset. 
```{r}
pacman::p_load(tidyverse)
pacman::p_load(bestNormalize)
```
\
## Loading of the radiomics dataset using the `readr` package
The radiomics dataset is loaded and assigned to a variable name **RDat**.
```{r}
library(readr)
RDat=read_csv("radiomics_completedata.csv", show_col_types = FALSE)
```
\
## Checking of null and missing values
```{r}
sum(is.na(RDat))
```
Based from the result 0, the dataset RDat has *no* null and missing values.
\
## Normality Test
Kolmogorov-Smirnov (`ks.test()`) test is used to check for normality of the dataset.
```{r, warning=FALSE}
RD1=RDat%>%select_if(is.numeric)
RD1=RD1[,-c(1:2)]
RD2=apply(RD1,2,function(x){ks.test(x,"pnorm")})
```
\
## Unlist the dataset RD2
The unlist produce a vector which containd all the atomic components which occur in RD2 dataset
```{r}
KS_list=unlist(lapply(RD2, function(x) x$p.value))
```
\
## Checking the number of variables that are not normally distributed
Counting the number of variables that are not normally distributed.
```{r}
sum(KS_list<0.05)
```
From the result, there are 428 variables that are not normally distributed. 
\
## Checking the number of variables that are normally distributed 
```{r}
sum(KS_list>0.05)
```
The result of 0 means that there is no nromally distributed variable. 
\
## Checking the variable with the maximum value p-value in the list
```{r}
which.max(KS_list)
```
From the result, the variable Kurtosis_hist.PET has the maximum p.value.
\
## Normalization of the dataset and checking of normality
The `orderNorm` is used for normalization. The Kolmogorov-Smirnov test is used for checking the normality of the dataset.
```{r,warning=F}
tempDFR=RDat[,c(3,5:length(names(RDat)))]
tempDFR=apply(tempDFR,2,orderNorm)
tempDFR=lapply(tempDFR, function(x) x$x.t)
tempDFR=tempDFR%>%as.data.frame()
testRD=apply(tempDFR,2,function(x){ks.test(x,"pnorm")})
testRD=unlist(lapply(testRD, function(x) x$p.value))
```
\
## Checking the number of variables which are normally distributed
```{r,warning=F}
sum(testRD>0.05)
```
From the result, there are 428 variables which are normally normally distributed. 
\
## Checking the number of variables of which are not normally distributed
```{r,warning=F}
sum(testRD<0.05)
```
From the result of the chunk, there is no more variable that is not normally distributed. 
\
## Collecting all the variables into one dataset
```{r}
RDat[,c(3,5:length(names(RDat)))]=tempDFR
```
\
## Checking for correlation
```{r}
CorMatrix=cor(RDat[,-c(1,2)])
heatmap(CorMatrix,Rowv=NA,Colv=NA,scale="none",revC = T)
```
\
## Transforming a some variables as categorical
```{r}
RDat$Institution=as.factor(RDat$Institution)
RDat$Failure.binary=as.factor(RDat$Failure.binary)
```
\
## Saving a normalize dataset as `normalRad`
```{r}
write.csv(RDat, "D:/FilesWorkOn&Saved/PhDStat/@MSUIIT/SY20222023/01FirstSemester/STT225_StatisticalComputing1/04FinalProject/02Dataset/FData/normalRad.csv", row.names=FALSE)
```
\
## Splitting of dataset into a training data and testing data
```{r}
splitter <- sample(1:nrow(RDat), round(nrow(RDat) * 0.8))
trainRDat <- RDat[splitter, ]
testRDat  <- RDat[-splitter, ]
```



