---
title: "Model 1 Support Vector Machine"
author: "Noel C. Sieras"
date: "2022-12-16"
output:
  pdf_document: default
  html_document: default
---
\
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'D:/FilesWorkOn&Saved/PhDStat/@MSUIIT/SY20222023/01FirstSemester/STT225_StatisticalComputing1/04FinalProject/02Dataset')
```
\
## Packages
```{r, echo=FALSE}
# Helper packages
library(dplyr)    # for data wrangling
library(ggplot2)  # for awesome graphics
library(rsample)  # for data splitting
library(readr)    #load dataset

# Modeling packages
library(caret)    # for classification and regression training
library(kernlab)  # for fitting SVMs
library(modeldata) #for Failure.binary data
library(forcats)
library(pdp)      # for partial dependence plots, etc.
library(vip)      # for variable importance plots
```
\
# MODEL 1 SVM

   Support Vector Machine (SVM) is a supervised learning technique that analyzes data and isolates patterns applicable to both classification and regression. The classifier is useful for choosing between two or more possible outcomes that depend on continuous or categorical predictor variables. Based on training and sample classification data, the SVM algorithm assigns the target data into any one of the given categories. The data is represented as points in space and categories are mapped in both linear and non-linear ways.
\
# LOADING OF REPROCESSED DATASET

We used the reprocessed data of radiomics_complete.csv (*normalRad.csv*) in performing support vectore machine.

Radiomics Dataset 197 Rows (Observations) of 431 Columns (Variables)
Failure.binary: binary property to predict
\
## Normalized Dataset
```{r}

radiomicsdat<- read_csv("normalRad.csv")
#View(radiomicsdt)
head(radiomicsdat)

# Load Failure.binary data

radiomicsdat$Failure.binary=as.factor(radiomicsdat$Failure.binary)
```
\
# Create training (80%) and test (20%) sets
```{r}
set.seed(123)  # for reproducibility

churn_split <- initial_split(radiomicsdat, prop = 0.8, strata = "Failure.binary")
split_train <- training(churn_split)
split_test  <- testing(churn_split)
```
\
## Classifiers
```{r}
# Linear (i.e., soft margin classifier)
caret::getModelInfo("svmLinear")$svmLinear$parameters

# Polynomial kernel
caret::getModelInfo("svmPoly")$svmPoly$parameters

# Radial basis kernel
caret::getModelInfo("svmRadial")$svmRadial$parameters
```
\
## Run SVM Model in the Training phase
Using **split_train**, we can tune an SVM model with radial basis kernel.
```{r}
set.seed(1854)  # for reproducibility
split_svm <- train(
  Failure.binary ~ ., 
  data = split_train,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  trControl = trainControl(method = "cv", number = 10),
  tuneLength = 10
)
```
\
## Plot and print SVM model with with radial basis kernel.
```{r}
# Plot results
ggplot(split_svm) + theme_light()

# Print results
split_svm$results
```
The accuracy of the model is **80.19%**
\
## Control and train of parameters
```{r}
class.weights = c("No" = 1, "Yes" = 10)

# Control params for SVM
ctrl <- trainControl(
  method = "cv", 
  number = 10, 
  classProbs = TRUE,                 
  summaryFunction = twoClassSummary  # also needed for AUC/ROC
)

split_train$Failure.binary=fct_recode(split_train$Failure.binary,No="0",Yes="1")

```

# Print the AUC values during Training
```{r}
# Tune an SVM
set.seed(5628)  # for reproducibility
train_svm_auc <- train(
  Failure.binary ~ ., 
  data = split_train,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  metric = "ROC",  # area under ROC curve (AUC)       
  trControl = ctrl,
  tuneLength = 10
)

# Print results
train_svm_auc$results
confusionMatrix(train_svm_auc)
```
On the average, the accuracy of the model is **83.44%**.
\
## Print the Top 20 important features during Training
```{r}
prob_yes <- function(object, newdata) {
  predict(object, newdata = newdata, type = "prob")[, "Yes"]
}

# Variable importance plot
set.seed(2827)  # for reproducibility
vip(train_svm_auc, method = "permute", nsim = 20, train = split_train, 
    target = "Failure.binary", metric = "auc", reference_class = "Yes", 
    pred_wrapper = prob_yes)
```
\
## Print the AUC values during Testing
\
```{r}
split_test$Failure.binary=fct_recode(split_test$Failure.binary,No="0",Yes="1")

# Tune an SVM with radial 
set.seed(5628)  # for reproducibility
test_svm_auc <- train(
  Failure.binary ~ ., 
  data = split_test,
  method = "svmRadial",               
  preProcess = c("center", "scale"),  
  metric = "ROC",  # area under ROC curve (AUC)       
  trControl = ctrl,
  tuneLength = 10
)

# Print results
test_svm_auc$results
confusionMatrix(test_svm_auc)
```
On the average, the accuracy of the model using the testing data is **62.5%**.


